### 2018-10-31 paper/complexity.tex
New version of section Complexity Considerations completed.

In this section, we calculate upper bounds for the total number of test
executions to be performed when testing for failures refinement 
and---as a corollary---trace refinement.
Theorem~\ref{th:failurestest} specifies that all tests $U_F(j),\ 0\le j < pq$
need to be executed, where $p$ denotes the number of nodes in the transition
graph of the reference process $P$, and $q\ge p$ is an estimate for the
maximal number of nodes in the SUT's transition graph. Therefore, we will first
calculate a bound for the number of test executions to be performed for test
$U_F(j)$ and then summarise  these bounds over all $j$ from $0$ to $pq-1$.

For the worst-case estimate, we first introduce a CSP reference process $\pmax$
which 
turns out to be--given a fixed alphabet $\Sigma$---the test 
model leading to the maximal number 
of test executions for every $U_F(j)$ when considering large alphabets. In the case
of small alphabets and large values of $pq-1$, it will turn out below that a variant
of $\pmax$ will lead to the maximal number of executions. The conditions for this
can also be clearly specified.

@@

% -------------------------------------------------------------------------
\subsection{A Reference Process With Maximal Test Complexity}

% -----------------------------------------------------------------------------
\subsubsection*{Process Definition and Basic Properties}
Given an alphabet $\Sigma$ of size $\card{\Sigma} = n\ge 2$, define
a collection of subsets of $\Sigma$ by
\begin{equation}\label{eq:defC}
{\cal C} = \{ A\subseteq\Sigma~|~\card{A} = n - \lfloor\frac{n}{2}\rfloor + 1 \}.
\end{equation}
%
With this choice of ${\cal C}$, define
\begin{equation}\label{eq:pmax}
\pmax = \Intchoice_{A\in{\cal C}} e:A\then \pmax
\end{equation}
%
The important properties of $\pmax$ are summarised in the following lemma.
%
\begin{lemma}\label{lemma:pmax}
Given alphabet $\Sigma$ with cardinality $\card{\Sigma} = n\ge 2$,
process $\pmax$ fulfils
%
\begin{eqnarray}
{}[\pmax/s]^0 & = & \Sigma \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxa}
\\
\trc(\pmax) & = & \Sigma^*
\label{eq:pmaxb}
\\
\minaccs(P/s) & = & {\cal C} \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxc}
\\
\minhits(P/s) & = & \minhits({\cal C}) \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxe}
\\
\card{\minhits(P/s)} & = & \binom{n}{\lfloor\frac{n}{2}\rfloor} 
\quad \text{for all $s\in \Sigma^*$}
\label{eq:pmaxd}
\\
\minhits({\cal C})  & = & \{ H\subseteq \Sigma~|~\card{H} = \lfloor\frac{n}{2}\rfloor\}
\label{eq:pmaxf}
\end{eqnarray}
\end{lemma}
\begin{proof}
 Since  
$
\bigcup_{A\in{\cal C}} A = \Sigma
$
by construction of ${\cal C}$, $[\pmax]^0 = \Sigma$. Since $\pmax/e = \pmax$ for all
$e\in\Sigma$, this proves statements (\ref{eq:pmaxa}) and (\ref{eq:pmaxb}).
The internal choice construct used in the specification of $\pmax$ implies
$\minaccs(\pmax) = {\cal C}$. Again, $\pmax/e = \pmax$ for all
$e\in\Sigma$ implies $\minaccs(\pmax/s) = {\cal C}$ for all traces of $\pmax$, so this shows (\ref{eq:pmaxc}). Statement (\ref{eq:pmaxe}) is a direct consequence
of (\ref{eq:pmaxc}).

@@

The number of executions specified in (\ref{eq:maxexec}) is indeed maximal for
all reference processes $P$ fulfilling $\minhits(P/s)\neq\varnothing$ for all
traces $s$. All these processes can never enter branch (\ref{eq:ufb}), and,
if an execution with an SUT entered branch (\ref{eq:ufa}), this would only lead to early termination of the whole test suite, because a failure has been detected.
As a consequence, $\card{\Sigma}^{j}$ is the maximal number of traces to be executed 
up to length $j$, and from Theorem~\ref{th:sperner} we know that the number
$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ of hitting sets to be tested at the end of each trace of length $j$ is already maximal.

Summing up  formula 
(\ref{eq:maxexec}) over all test cases $U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum of a geometric progression,  this results in
%
\begin{equation}\label{eq:pmaxcomplexity}
\sum_{j=0}^{pq-1} n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}  = 
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}
\end{equation}
%
as the maximal number of test executions to be performed when testing an error-free
SUT $Q$ against reference process $\pmax$. This maximum is met when $\trc(Q) = \Sigma^*$.

@@

\item For $j = \#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has 
$(\card{\minhits(Z/\ol s)} - 1)$ fewer executions than $(Z\parallel[\Sigma] U_F(j))$, because 
$Z(\ol s,\ol s)$ has an empty set of hitting sets, since 
$Z(\ol s,\ol s) = (Z/\ol s)\intchoice\Stop$. As a consequence, test case
${\ol U}_F(\#\ol s)$ can only enter branches 
(\ref{eq:ufa})  or (\ref{eq:zcont}) after having run through trace
$\ol s$. The former branch is available if and only if it is available in $U_F(\#\ol s)$, the latter just stops the test after issuing the $\epass$ event.

@@

The construction of $Z(\ol s)$ can be generalised by using $k$ traces 
${\ol s}_i, \ i = 1\dots,k$ of the same length $\ol j$, 
after that the modified process behaves as 
$(Z/{\ol s}_i)\intchoice \Stop$. Generalising the case distinction above, 
\begin{enumerate}
\item the number of test executions remains unchanged for $j < \ol j$,
\item the number of   executions lost for $j = \ol j$ is at least 
$k$ (assuming that for all $i$, $\card{\minhits(Z/{\ol s}_i)} = 1$), and
\item the number of executions won for $j > \ol j$ is $k$.
\end{enumerate}
%
The number $k$ of traces ${\ol s}_i,\ i=1,\dots,k$ is bounded by the maximal number
of traces that can be executed by $(Z\parallel[\Sigma] U_F(j))$, which is bounded by
$\card{\Sigma}^{\ol j}$ (this bound is reached, for example, 
for process $\pmax$ discussed above). If $k$ equals this maximal number, there is no need
to execute test case ${\ol U}_F(\ol j)$ any longer, because it can only stop after $\epass$, whenever a trace of length $\ol j$ is executed.

Summing up over all test cases for such a $k \le \card{\Sigma}^{\ol j}$, this results in 
an upper bound of
\begin{equation}
 \card{\Sigma}^{\ol j}(pq - \ol j      - 1)
\end{equation}
additional test executions in comparison to the number of executions needed for all tests
of $Z$. Now we can apply this construction to all traces lengths from $0$ to $pq-2$ (we have 
explained above that test executions are lost for trace lengths greater than $pq-2$). This results in an upper bound of
\begin{equation}
\sum_{j=0}^{pq-2}  n^j(pq - j      - 1) =  
\frac{n^{pq} + pq - npq -1}{(n - 1)^2}
\quad\text{with $n = \card{\Sigma}$}
\end{equation}
additional test executions that can be gained by applying the construction technique
explained above for $Z(\ol s)$, extended to all traces of length less or equal to $pq-2$.
%
Summarising this result and the one previously obtained in (\ref{eq:maxexec}),
the following complexity bounds are obtained.
%
\begin{theorem}
Given a process alphabet $\Sigma$ with cardinality $
\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
the normalised transition graph of $P$ has $p$ states, and the fault domain 
${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then the number of test executions to be performed   using the complete test 
suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} has the following bounds.
%
\begin{enumerate}
\item If $pq \le \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
test suite  
requires at most 
%
\begin{equation*}
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}
\end{equation*}
%
test executions. This upper bound is reached for process $\pmax$ specified above.

\item If $pq > \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
test suite  
requires less than 
\begin{equation*}
pq\frac{1-n^{pq}}{1-n} +
\frac{n^{pq} + pq - npq -1}{(n - 1)^2} \qquad\text{@todo}
\end{equation*}
executions.
\end{enumerate}
\xbox
\end{theorem}

% -----------------------------------------------------------------------------
\subsubsection*{Considering Empty Collections of Minimal Hitting Sets -- Old version}

In the following specification of $Z$, given traces $s_1,s_2\in\Sigma^*$, the Boolean 
expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$. Expression $s_1\le s_2$ holds if and only if $ s_1 < s_2 \vee s_1 = s_2$ is true.
\begin{eqnarray*}
Z & = & Z_1(\varepsilon)
\\
Z_1(s) & = & (s < \ol s)\&\big( \Intchoice_{A\in{\cal C}} e:A\then Z_1(s.e)  \big)
\\ & & \extchoice
\\ & & (s \not\le \ol s) \& \pmax
\\ & & \extchoice
\\ & & (s = \ol s) \& \big( (\Intchoice_{A\in{\cal C}} e:A\then \pmax)  
\intchoice \Stop\big)
\end{eqnarray*}
%
It is easy to see that $\trc(Z) = \trc(\pmax) =\Sigma^*$ and that for all 
traces $s\neq\ol s$, $\minhits(Z/s) = \minhits(P/s) = \minhits({\cal C})$ as specified
in Lemma~\ref{lemma:pmax}.

When comparing the test cases $U_F(j)$ generated from $\pmax$ and the test cases
$U_F^Z(j)$ created from $Z$, we note that
\begin{enumerate}
\item For $j < \#\ol s$, the test cases  $U_F(j)$ and $U_F^Z(j)$ are identical.
\item For $j = \#\ol s$, branch (\ref{eq:ufb}) can be entered by $U_F^Z(j)/\ol s$, while it is blocked for $U_F(j)/\ol s$, and branch (\ref{eq:ufd}) is blocked
for  $U_F^Z(j)/\ol s$, while it can be entered by $U_F(j)/\ol s$. As a consequence,
the maximal number of executions that can be performed by $U_F^Z(j)$ against the SUT 
is 
\begin{equation}
(n^{j} - 1)\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor} + 1
\end{equation}
according to (\ref{eq:maxexec}): we win one additional execution by entering 
branch (\ref{eq:ufb}) after $\ol s$ and lose $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ executions
by no longer being able to enter branch (\ref{eq:ufd}) after $\ol s$.

\item For $j > \#\ol s$, one execution is won by entering 
branch (\ref{eq:ufb}) after $\ol s$, no executions are lost.
\end{enumerate}
%
Summing up over $j = 0,\dots,pq-1$, this results in a difference of 
%
\begin{equation}
(pq  - \#\ol s) - \binom{n}{\lfloor\frac{n}{2}\rfloor}
\end{equation}
executions for the test cases $U_F^Z(j),\ j=0,\dots,pq-1$ in comparison to 
the test cases $U_F(j),\ j=0,\dots,pq-1$. To assess this formula, see the function
table for the first 19  arguments of $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ shown
in Table~\ref{table:binom}.
%
\begin{table}[htp]
\caption{Function table of $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ for arguments
$n = 2,\dots,20$.}
\begin{center}
\begin{tabular}{c||c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c}
$n$ & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13  & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\\hline
$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ &
     2 & 3 & 6& 10 & 20 & 35 & 70 & 126 & 252 &
     462&924&1716&3432&6435&12870&24310&48620&92378&184756
\end{tabular}
\end{center}
\label{table:binom}
\end{table}
%
The analysis above leads to the following complexity result.
%
\begin{theorem}
Given a process alphabet $\Sigma$ with cardinality $
\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
the normalised transition graph of $P$ has $p$ states, and the fault domain 
${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then, if $pq < \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
test suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ according to Theorem~\ref{th:failurestest}
requires at most 
%
\begin{equation*}
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}
\end{equation*}
%
test executions.
\xbox
\end{theorem}

% -------------------------------------------------------------------------
\subsection{Estimation of $\ell$}
The first factor $\ell$ has worst-case upper bound $\ell\le \card{\Sigma}^j$.
As an example where this upper bound is really met, we consider the reference
process
\[
RUN(\Sigma) = e:\Sigma \then RUN(\Sigma).
\]
The normalised transition graph of this process has a single state, and its initials
are $[RUN(\Sigma)]^0 = \Sigma$. Therefore, the associated test process $U_F(j)$ can
never enter branches (\ref{eq:ufa}) and (\ref{eq:ufb}), but there are exactly
$\card{\Sigma}^j$ different traces of length $j$ exercising branch (\ref{eq:ufc})
for each of their events.

% -------------------------------------------------------------------------
\subsection{Estimation of $h$}
Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
acceptances, the cardinality   $h = \card{\minhits(P/s)}$ has the upper bound
\[
h \le \binom{n}{\lfloor\frac{n}{2}\rfloor}, \quad \text{where $n = \card{\Sigma}$.}
\]
This follows from Theorem~\ref{th:sperner}, since $\minhits(P/s)$ is a
Sperner Family~(see Section~\ref{sec:sperner}). The next theorem shows that
this upper bound can really be reached by collections of minimal hitting sets
associated with a CSP process state.

\begin{theorem}
\label{th:upperboundh}
Let  $\Sigma$  be an alphabet with cardinality $n\ge 2$.  Then
there exists a CSP process $P$ with
\[
\card{\text{minHit}(P)} = \binom{n}{\lfloor\frac{n}{2}\rfloor}.
\]
\end{theorem}
\begin{proof}
Let ${\cal C}$ be the collection of all subsets of $\Sigma$ with cardinality
$n-\lfloor{\frac{n}{2}}\rfloor+1$. Then the CSP process
\[
P = \Intchoice_{A\in {\cal C}} e:A\then P
\]
fulfils $\minaccs(P) = {\cal C}$. Let $H$ be any minimal hitting set of ${\cal C}$. Then
$H$ contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because
otherwise $\card{\Sigma\setminus H} > n-\lfloor{\frac{n}{2}}\rfloor$, and any
subset $A\subseteq \Sigma\setminus H$ with cardinality
$n-\lfloor{\frac{n}{2}}\rfloor+1$  would be contained in ${\cal C}$, but satisfy
$A\cap H=\varnothing$. Since
$\lfloor{\frac{n}{2}}\rfloor+n-\lfloor{\frac{n}{2}}\rfloor+1=n+1$, we
conclude
 that any $\lfloor{\frac{n}{2}}\rfloor$-element subset of $\Sigma$
 intersects  every element of ${\cal C}$.  Therefore, every minimal hitting set of ${\cal C}$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements,
so $\card{\minhits({\cal C})} = \binom{n}{\lfloor{\frac{n}{2}}\rfloor}$.
\xbox
\end{proof}
%
To get an approximation of the maximal size of $\minhits(P/s)$ for large $n$,
recall Stirling's approximation~\cite[p.~112]{Graham:1994:CMF:562056}
\[
m! \approx \sqrt{2\pi m} \cdot \big( \frac{m}{e} \big)^m, \quad\text{where $e$ denotes the Euler Number.}
\]
Applying this approximation when $n = \card{\Sigma}$ is even and $m = (n/2)$
results in
\[
h \le \binom{2m}{m} \approx \frac{4^m}{\sqrt{\pi m}} =
\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}.
\]
%
We note that, due to this approximation result,
$h$ is   smaller than $2^{\card{\Sigma}}-1$, the cardinality of the
non-empty subsets of $\Sigma$.

\begin{theorem}\label{th:numexecP}
Let $U_F(j)$ be a test case derived from the 
process $P$ specified in the proof of Theorem~\ref{th:upperboundh}. Then
$U_F(j)$ has
\[
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot \card{\Sigma}^j
\]
executions.
\end{theorem}
\begin{proof}
Since the union of all minimal acceptances of $P$ equals $\Sigma$, 
the initials of $P$ satisfy $[P]^0 = \Sigma$. By construction, $P/e = P$
for any $e\in\Sigma$, so $\trc(P) = \Sigma^*$ and $[P/s]^0 = \Sigma$
for all $s\in\Sigma^*$. As a consequence, $U_F(j)$
may execute $\card{\Sigma}^j$ different traces, each time executing branch (\ref{eq:ufc}),
before finally entering branch (\ref{eq:ufd}). As shown in Theorem~\ref{th:upperboundh},
$\card{\minhits(P/s)} =  \binom{n}{\lfloor\frac{n}{2}\rfloor}$, so there are
$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ possible executions of branch (\ref{eq:ufd})
after every trace of length $j$. This proves the theorem.
\xbox
\end{proof}

% -------------------------------------------------------------------------
\subsection{Upper Bounds of Test Executions for Checking Failures Refinement}

By Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(j)$ for
$j = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
we use the formula for the sum of the geometric progression.
%
\[
h\cdot \big( \frac{1-\card{\Sigma}^{pq}}{1-\card{\Sigma}} \big),\
 \text{or, asymptotically,}\  O(\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}\cdot\card{\Sigma}^{(pq-1)}).
 \]
%
From Theorem~\ref{th:upperboundh} we know that the worst-case bound for $h$
above cannot be further reduced, since the full collection of minimal hitting
sets needs to be checked at the end of each execution of $U_F(j)$.

In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
non-empty subset of $\Sigma$ whose events cannot be completely refused in a
given process state of the reference model; this leads to a worst-case
estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
offered to the SUT in the last step of the test execution, so our approach
reduces the number of test executions in comparison
to~\cite{Hennessy:1988:ATP:50497} by a factor of $1/{\sqrt{2\pi
\card{\Sigma}}}$. In Fig.~\ref{fig:minhita}, the reduction is visualised by a
function plot of $2^{\card{\Sigma}}$ versus
$\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.

In~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors also use minimal hitting
sets\footnote{However, they are denoted by {\it minimal acceptances}
in~\cite{DBLP:conf/icfem/CavalcantiG07}.}, but they do not give an upper
bound for the number of test executions to be performed.

% .....................................................................................
 \begin{figure}
 %%\hspace*{-40mm}
 \begin{center}
\includegraphics[width=.8\textwidth]{minhit-fig.pdf}
\end{center}
\vspace*{-10mm}
\caption{Function plot $2^{\card{\Sigma}}$ versus $\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.}
 \label{fig:minhita}
 \end{figure}
% .......................................................................................

% -------------------------------------------------------------------------
\subsection{Upper Bounds of Test Executions for Checking Trace Refinement}

According to Theorem~\ref{th:tracetest}, a complete test suite checking trace
refinement just contains the adaptive test case $U_T(pq-1)$. As derived for
$U_F(j)$ above, the number of executions performed by $(Q\parallel[\Sigma]
U_T(pq-1))$ is bounded by $\card{\Sigma}^{pq-1}$.

>>>>

%In this section, we calculate estimates for the maximal number of test
%executions to be performed when testing for failures refinement 
%and---as a corollary---trace refinement.
%Theorem~\ref{th:failurestest} specifies that all tests $U_F(j),\ 0\le j < pq$
%need to be executed, where $p$ denotes the number of nodes in the transition
%graph of the reference process $P$, and $q\ge p$ is an estimate for the
%maximal number of nodes in the SUT's transition graph. Therefore, we will first
%calculate a bound for the number of test executions to be performed for test
%$U_F(j)$ and then summarise  these bounds over all $j$ from $0$ to $pq-1$.
%
%For the worst-case estimate, we first introduce a CSP reference process $\pmax$
%which 
%turns out to be--given a fixed alphabet $\Sigma$---the test 
%model leading to the maximal number 
%of test executions for every $U_F(j)$ when considering large alphabets. In the case
%of small alphabets and large values of $pq-1$, it will turn out below that a variant
%of $\pmax$ will lead to the maximal number of executions. The conditions for this
%can also be clearly specified.

@@

% -------------------------------------------------------------------------
\subsection{Estimates for the Maximal Number of Failures Test Executions}

% -----------------------------------------------------------------------------
\subsubsection*{A Reference Process}
Given an alphabet $\Sigma$ of size $\card{\Sigma} = n\ge 2$, define
a collection of subsets of $\Sigma$ by
\begin{equation}\label{eq:defC}
{\cal C} = \{ A\subseteq\Sigma~|~\card{A} = n - \lfloor\frac{n}{2}\rfloor + 1 \}.
\end{equation}
%
With this choice of ${\cal C}$, define
\begin{equation}\label{eq:pmax}
\pmax = \Intchoice_{A\in{\cal C}} e:A\then \pmax
\end{equation}
%
The relevant properties of $\pmax$ are summarised in the following lemma.
%
\begin{lemma}\label{lemma:pmax}
Given alphabet $\Sigma$ with cardinality $\card{\Sigma} = n\ge 2$,
process $\pmax$ fulfils
%
\begin{eqnarray}
{}[\pmax/s]^0 & = & \Sigma \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxa}
\\
\trc(\pmax) & = & \Sigma^*
\label{eq:pmaxb}
\\
\minaccs(P/s) & = & {\cal C} \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxc}
\\
\minhits(P/s) & = & \minhits({\cal C}) \quad\text{for all $s\in \Sigma^*$}
\label{eq:pmaxe}
\\
\card{\minhits(P/s)} & = & \binom{n}{\lfloor\frac{n}{2}\rfloor} 
\quad \text{for all $s\in \Sigma^*$}
\label{eq:pmaxd}
\\
\minhits({\cal C})  & = & \{ H\subseteq \Sigma~|~\card{H} = \lfloor\frac{n}{2}\rfloor\}
\label{eq:pmaxf}
\end{eqnarray}
\end{lemma}
\begin{proof}
 Since  
$
\bigcup_{A\in{\cal C}} A = \Sigma
$
by construction of ${\cal C}$, $[\pmax]^0 = \Sigma$. Since $\pmax/e = \pmax$ for all
$e\in\Sigma$, this proves statements (\ref{eq:pmaxa}) and (\ref{eq:pmaxb}).
The internal choice construct used in the specification of $\pmax$ implies
$\minaccs(\pmax) = {\cal C}$. Again, $\pmax/e = \pmax$ for all
$e\in\Sigma$ implies $\minaccs(\pmax/s) = {\cal C}$ for all traces of $\pmax$, so this shows (\ref{eq:pmaxc}). Statement (\ref{eq:pmaxe}) is a direct consequence
of (\ref{eq:pmaxc}).

@@

The number of executions specified in (\ref{eq:maxexec}) is indeed maximal for
all reference processes $P$ fulfilling $\minhits(P/s)\neq\varnothing$ for all
traces $s$: All these processes can never enter branch (\ref{eq:ufb}), and,
if an execution with an SUT entered branch (\ref{eq:ufa}), this would only lead to early termination of the whole test suite, because a failure has been detected.
As a consequence, $\card{\Sigma}^{j}$ is the maximal number of traces to be executed 
up to length $j$, and from Theorem~\ref{th:sperner} we know that the number
$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ of hitting sets to be tested at the end of each trace of length $j$ is already maximal.

Summing up  formula 
(\ref{eq:maxexec}) over all test cases $U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum of a geometric progression,  this results in
%
\begin{equation}\label{eq:pmaxcomplexity}
\sum_{j=0}^{pq-1} n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}  = 
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}
\end{equation}
%
as the maximal number of test executions to be performed when testing an error-free
SUT $Q$ against reference process $\pmax$. This maximum is met when $\trc(Q) = \Sigma^*$.
If one is only interested in the order of magnitude, the maximal nmber of executions 
may be re-written as 
%
\begin{equation} 
\label{eq:maxO}
O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
\end{equation}

@@

\item For $j = \#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has 
$(\card{\minhits(Z/\ol s)} - 1)$ fewer executions than $(Z\parallel[\Sigma] U_F(j))$, because 
$Z(\ol s,\ol s)$ has an empty set of hitting sets, since 
$Z(\ol s,\ol s) = (Z/\ol s)\intchoice\Stop$. As a consequence, test case
${\ol U}_F(\#\ol s)$ can only enter branches 
(\ref{eq:ufa})  or (\ref{eq:ufb}) after having run through trace
$\ol s$. The former branch is available if and only if it is available in $U_F(\#\ol s)$, the latter just stops the test after issuing the $\epass$ event.

@@

The construction of $Z(\ol s)$ can be generalised by using $k$ traces 
${\ol s}_i, \ i = 1\dots,k$ of the same length $\ol j$, 
after that the modified process behaves as 
$(Z/{\ol s}_i)\intchoice \Stop$. Generalising the case distinction above, 
\begin{enumerate}
\item the number of test executions remains unchanged for $j < \ol j$,
\item the number of   executions lost for $j = \ol j$ is at least 
$k$ (assuming that for all $i$, $\card{\minhits(Z/{\ol s}_i)} = 1$), and
\item the number of executions won for $j > \ol j$ is $k$.
\end{enumerate}
%
The number $k$ of traces ${\ol s}_i,\ i=1,\dots,k$ with fixed length $\#{\ol s}_i=\ol j$ is bounded by the maximal number
of traces that can be executed by $(Z\parallel[\Sigma] U_F(\ol j))$, which is bounded by
$\card{\Sigma}^{\ol j}$. If $k$ equals this maximal number, there is no need
to execute test case ${\ol U}_F(\ol j)$ any longer, because it can only stop after $\epass$, whenever a trace of length $\ol j$ is executed.

Summing up over all test cases ${\ol U}_F(j),\ j = \ol j,\dots,pq-1$, for a given 
 $k \le \card{\Sigma}^{\ol j}$, this results in 
an upper bound of
\begin{equation}
 \card{\Sigma}^{\ol j}(pq - \ol j      - 1)
\end{equation}
additional test executions in comparison to the number of executions needed for all tests
of $Z$. The maximal value of this term is reached for $\ol j = pq - 2$: this is the maximal exponent where $(pq - \ol j      - 1)$ is still greater than zero. We conclude that the
maximal increase of test executions to be gained if processes possess failures $(s,\Sigma)$
is of the order $\card{\Sigma}^{pq - 2}$. This can be neglected in comparison to 
the worst case applicable for processes that do not possess any failures $(s,\Sigma)$, because 
formula (\ref{eq:maxO}) shows that the number of test executions required for the latter   is of order $\card{\Sigma}^{pq - 1}$.

These considerations lead to the following theorem
%
\begin{theorem}
Given a process alphabet $\Sigma$ with cardinality $
\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
the normalised transition graph of $P$ has $p$ states, and the fault domain 
${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then the maximal 
number of test executions to be performed   using the complete test 
suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} is of order
%
\begin{equation*} 
O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
\end{equation*} 
\xbox
\end{theorem}
%
In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
non-empty subset of $\Sigma$ whose events cannot be completely refused in a
given process state of the reference model; this leads to a worst-case
estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
offered to the SUT in the last step of the test execution. This number is significantly larger than the worst-case estimate $\binom{n}{\lfloor n/2\rfloor}$ calculated above for the 
hitting sets to be checked.
In Fig.~\ref{fig:minhita}, the reduction is visualised by plots of
the two functions.

In~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors also use minimal hitting
sets\footnote{However, they are denoted by {\it minimal acceptances}
in~\cite{DBLP:conf/icfem/CavalcantiG07}.}, but they do not give an estimate for the number of test executions to be performed.

% .....................................................................................
 \begin{figure}
 %%\hspace*{-40mm}
 \begin{center}
\includegraphics[width=.8\textwidth]{curvecomparison.pdf}
\end{center}
%\vspace*{-10mm}
\caption{Function plot $2^{\card{\Sigma}}$ versus $\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.}
 \label{fig:minhita}
 \end{figure}
% .......................................................................................

% -------------------------------------------------------------------------
\subsection{Estimates for the Maximal Number of Trace Test Executions}

According to Theorem~\ref{th:tracetest}, a complete test suite checking trace
refinement just contains the adaptive test case $U_T(pq-1)$. As derived for
$U_F(j)$ above, the maximal number of executions to be performed by $(Q\parallel[\Sigma]
U_T(pq-1))$ is of order  $O\big(\card{\Sigma}^{pq-1}\big)$.
